Individual heterogeneity in the functional network topography of the medial prefrontal cortex

### 2019-05-01 Manuscript.Rmd
new version of similarities with other work

  A useful way to characterize individual-specific brain organization is to examine patterns of resting-state functional connectivity (rsFC; i.e. BOLD time-series correlations between pairs of locations in the brain). Connectome-based analyses of rsFC have been fruitful in characterizing individualized functional topologies that correspond well to task-induced activity [@Gordon2017; @Laumann2015; @Smith2009; @Tobyne2017]. A functional connectome can be represented in the form of a network, and graph theoretic methods can be applied to analyze the network's structure [@Bassett2017; @Rubinov2010]. In the context of network analysis, community detection algorithms subdivide brain networks into sets of nodes that share more connections with each other than with the rest of the network [@Fortunato2016; @Garcia2017]. Here we use the technique of spectral partitioning (SP), an efficient community detection algorithm that deterministically subdivides a network into two communities [@Belkin2003; @Chung1999; @Fiedler1975]. SP has been used to characterize the posterior-anterior functional gradient of the insula using resting-state fMRI (rsfMRI) data [@Tian2018], and was shown to robustly and reliably separate both simulated and actual primate ECoG networks [@Toker2019]. We use SP here to identify subsets of nodes within mPFC and PCC that share spontaneously covarying temporal activation patterns during rest.

@@

Communities (i.e. clusters) were identified using the SP algorithm [@Belkin2003; @Chung1999; @Fiedler1975; @Higham2007]. First, each network was represented as an $n$ x $n$ network weight matrix $W$ as described above (where $n$ equals the number of vertices in the search space, 4,801). This matrix was then transformed into its symmetric normalized Laplacian form

@@

The thresholded partitions in our work qualitatively resembled the topographical arrangement of DN sub-networks A and B proposed by Braga and Buckner [-@Braga2017]. We explored this similarity by reproducing their seed-based connectivity approach in three of our subjects. In their original work, Braga and Buckner manually selected individual vertices in areas including the dorsolateral prefrontal cortex (DLPFC), temporo-parietal junction (TPJ), and parahippocampal cortex (PHC) that produced two spatially anticorrelated networks [@Braga2017; @Braga2019]. Therefore, we hypothesized that if the communities we identified on the medial wall matched their proposed networks, we should be able to reproduce our partitionings by selecting seeds in these diagnostic areas. We confirmed seed placement based on functional connectivity patterns observed in TPJ and PHC. The whole-brain seed-based functional connectivity maps for these subjects are juxtaposed with the corresponding community detection results in Figure 8.

Visual inspection of these networks shows a remarkable similarity between our DN community and the previously reported sub-network A. However, the non-DN community filled areas not covered by either DN-A or DN-B. A potential explanation is that SP fractionated the search space into communities that are maximally distant in function, whereas the sub-networks proposed by Braga and Buckner are still considered to be part of DN. Subthreshold vertices in the SP partitionings partially overlap with DN-B vertices, suggesting that this network could act as a dynamic intermediary between DN-A (the DN community) and other networks.

Understanding the interaction of these three networks requires future study. Nevertheless, these qualitative similarities are important, as they collectively suggest that canonical DN regions consist of at least two distinct communities. Results from the present investigation suggest that the topographic organization of the communities is variable across a large sample of participants, especially in mPFC, and that the SP algorithm is a useful and reliable tool for identifying these communities in individuals.

@@

```{r BandB comparison (Figure 8), out.width = "100%", fig.align = "center", fig.cap= "Qualitative comparison between maps produced by seed-based connectivity based on Braga and Buckner's (2019) criterion, and those produced by spectral partitioning. Networks A and B were produced by selecting seeds in diagnostic TPJ, DLPFC and PHC. In each case, our DN and non-DN networks map well to their proposed functional organization, albeit with some discrepancies on network A/non-DN."}

@@

>>>>

  A useful way to characterize individual-specific brain organization is to examine patterns of resting-state functional connectivity (rsFC; i.e. BOLD time-series correlations between pairs of locations in the brain). Connectome-based analyses of rsFC have been fruitful in characterizing individualized functional topologies that correspond well to task-induced activity [@Gordon2017; @Laumann2015; @Smith2009; @Tobyne2017]. A functional connectome can be represented in the form of a network, and graph theoretic methods can be applied to analyze the network's structure [@Bassett2017; @Rubinov2010]. In the context of network analysis, community detection algorithms subdivide brain networks into sets of nodes that share more connections with each other than with the rest of the network [@Fortunato2016; @Garcia2017]. Here we use the technique of spectral partitioning (SP), an efficient community detection algorithm that deterministically subdivides a network into two communities [@Belkin2003; @chung1997; @Fiedler1975]. SP has been used to characterize the posterior-anterior functional gradient of the insula using resting-state fMRI (rsfMRI) data [@Tian2018], and was shown to robustly and reliably separate both simulated and actual primate ECoG networks [@Toker2019]. We use SP here to identify subsets of nodes within mPFC and PCC that share spontaneously covarying temporal activation patterns during rest.

@@

Communities (i.e. clusters) were identified using the SP algorithm [@Belkin2003; @chung1997; @Fiedler1975; @Higham2007]. First, each network was represented as an $n$ x $n$ network weight matrix $W$ as described above (where $n$ equals the number of vertices in the search space, 4,801). This matrix was then transformed into its symmetric normalized Laplacian form

@@

The thresholded partitions in our work qualitatively resembled the topographical arrangement of DN sub-networks A and B proposed by Braga and Buckner [-@Braga2017]. We explored this similarity by reproducing their seed-based connectivity approach in three of our subjects. In their original work, Braga and Buckner manually selected individual vertices in areas including the dorsolateral prefrontal cortex (DLPFC), temporo-parietal junction (TPJ), and parahippocampal cortex (PHC) that produced two spatially anticorrelated networks [@Braga2017; @Braga2019]. Therefore, we hypothesized that if the communities we identified on the medial wall matched their proposed networks, we should be able to reproduce our partitionings by selecting seeds in these diagnostic areas. For functional connectomes from two individuals (100307 and 101006), we selected seeds for networks A and B in TPJ (seed indices), and confirmed their placement based on functional connectivity patterns observed in TPJ, PCC, DLPFC, and PHC. The whole-brain seed-based functional connectivity maps for two individuals are juxtaposed with the corresponding community detection results in Figure 8.

Visual inspection of these networks shows a remarkable similarity between our DN community and the previously reported sub-network A. However, the non-DN community filled areas not covered by either DN-A or DN-B, further evidencing its disentanglement from DN processes. Subthreshold vertices from the DN community overlaped with DN-B vertices, suggesting that this network could act as a dynamic intermediary between DN-A and other networks.

Understanding the interaction of these three networks will require future examination. Nevertheless, these qualitative similarities are important, as they collectively suggest that: (1) canonical DN regions can be topographically fractionated into DN and non-DN; (2) the DN community can be further partitioned into two sub-networks; (3) the SP algorithm is a useful and reliable tool for identifying this refined topography; and (4) the topographic organization of the communities is variable across a large sample of participants, especially in mPFC.

@@

```{r BandB comparison (Figure 8), out.width = "100%", fig.align = "center", fig.cap= "Qualitative comparison between DN sub-networks A and B from Braga et al. (2019) and SP communities for two individuals. Upper: networks A and B produced by selecting seeds in TPJ, with our search space delineated by black borders. Bottom: thresholded communities (indicated by borders) show strong resemblance between the DN community and network A. The non-DN community covers sections of cortex not associated with either DN sub-network."}

@@

\pagebreak
